%!TEX root = main-doc.tex
%
% File: background.tex
%
% Date: ? 
%
% Description:
%   The background is given to ...
%
%
\chapter{Background} \label{chap:background}

\summary{Give chapter summary}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Warehousing}
Information is incredibly important \cite{golfarelli:2009:dwd} 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Early Results}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Result Differentiation}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodologies}
In order to develop a model for extendible multi-dimensional sparse arrays, the current representations of extendible multi-dimensional arrays must be analysed. The study of the current representations will give a clear indication of how to integrate the representation of extendible multi-dimensional sparse arrays, so that the model can easily be integrated into the current system, preventing unnecessary additional costs. Once a model has been developed, compression of data as well as computing speed will be assessed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sparse Array Representations}
Hairong Wang \cite{wang:2014sar}
\begin{itemize}
	\item Space linear in the number of non-zero elements
	\item Acceleration of operations using General Purpose Computing on Graphics Processing Unit
	\item 2-dimensional arrays
	\begin{itemize}
		\item Compressed Row (or alternatively Column) Storage (CRS/CCS)
	\end{itemize}
	\item multi-dimensional sparse arrays
	\begin{itemize}
		\item Bit-Encoded Sparse Storage (BESS) 
		\item Extended the CRS/CCS known as xCRS and xCCS
		\item PATRICIA Trie Compressed Storage (PTCS)
	\end{itemize}
	\item New storage techniques
	\begin{itemize}
		\item Bit-Encoded Compressed Row (or Column) storage (BxCRS/BxCCS)
		\item Hybrid approach (Hybrid) that combines BESS with xCRS
	\end{itemize}
\end{itemize}
Matrix operations include: 
\begin{itemize}
	\item Get, Insert, Delete, Update
	\item Addition, Multiplication, Subtraction, Division
\end{itemize}
2D sparse array can be represented as a linked list or as triplicate 


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Dynamic Sparse Arrays}
Increasing Density
\begin{itemize}
	\item Bounds and structure remain the same
	\item new elements are slot into a chunk where there was a free space
\end{itemize}
Increasing Bound


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Big Data Warehouse}
\begin{itemize}
	\item Bill Inmon Defines: A data warehouse is a subject-oriented, non-volatile, integrated, time variant collection of data created for the purpose of management’s decision making.
	\begin{itemize}
		\item A big data solution is a technology and data warehousing is an architecture.
	\end{itemize}
	\item Ralph Kimball defines: A data warehouse is a copy of transaction data specifically structured for query and analysis.
	\item Volume - Scalability, storage to grow - investigate using a parallel file system
	\item Velocity - Cope with speed, rate at which data comes in
	\item Variety - Data types
	\item Veracity - validity
	\item Value
	\item Fault tolerance must be researched
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Big Data Representations}
\begin{itemize}
	\item Capable of holding very large amounts of data.
	\item Hold the data in inexpensive storage devices.
	\item Processing is done by the “Roman census” method.
	\item Data is stored in an unstructured format.
	\item Hierarchical Data Format (HDF5)
	\begin{itemize}
		\item Data model, library, and file format for storing and managing data
		\item Supports an unlimited variety of datatypes, and is designed for flexible and efficient I/O and for high volume and complex data. 
		\item Portable and extensible, allowing applications to evolve in their use of HDF5.
		\item Tools and applications for managing, manipulating, viewing, and analyzing data in the HDF5 format. 
		\item Organized in a hierarchical structure, with two primary structures: groups and datasets. 
		\begin{itemize}
			\item group: a grouping structure containing instances of zero or more groups or datasets, together with supporting metadata. 
			\item dataset: a multidimensional array of data elements, together with supporting metadata. 
		\end{itemize}
	\end{itemize}
	\item Hadoop
	\begin{itemize}
		\item Distributed storage and processing of very large data sets. 
		\item Consists of computer clusters built from commodity hardware.
		\item Assumption that hardware failures are common occurrences and should be automatically handled by the framework
		\item Hadoop can run parallel queries over flat files. This allows it do basic operational reporting on data in its original form.
		\item Hadoop splits files into large blocks and distributes them across nodes in a cluster. It then transfers packaged code into nodes to process the data in parallel
		\item Hadoop framework includes following four modules:
		\begin{itemize}
			\item Hadoop Common: These libraries provides filesystem and OS level abstractions and contains the necessary Java files and scripts required to start Hadoop.
			\item Hadoop YARN: This is a framework for job scheduling and cluster resource management.
			\item Hadoop Distributed File System (HDFS™): A distributed file system that provides high-throughput access to application data.
			\item Hadoop MapReduce: This is YARN-based system for parallel processing of large data sets.
		\end{itemize}
	\end{itemize}
	\item Network Common Data Form (NetCDF)
	\begin{itemize}
		\item Set of software libraries and self-describing, machine-independent data formats that support the creation, access, and sharing of array-oriented scientific data.
	\end{itemize}
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{TileDB storage manager}
\begin{itemize}
	\item Used for data which is represented as multi-dimensional arrays.
	\item Sparse arrays significantly impact storage and application performance
	\item when storing data cells that are accessed together should be co-located -  global cell orders in data tiles in sparse arrays include row-major vs column-major
	\item bookkeeping is very important for locating the correct data as exact distribution of the non-empty cells is unknown
	\item Still working on understanding the internals and basic operations in more depth
	\item Downloaded some files on reading/writing to sparse arrays (C) - DBexamples on Github - Intel HCS compiler - link between FPGA and GPU
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{FGPAs}
Integrated circuit consisting of an array of identical logic blocks with programmable interconnections.
-I do not feel that FGPAs would be the best suited for the research. \cite{roth:2016:dsd}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Impact}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{New Results}
